---
title: "Analysis of Coverage"
author: "Bernhard Konrad"
date: "December 22, 2015"
output:
  html_document:
    keep_md: true
---

In this document we run a data quality analysis to check the observed receiver coverage. This is to control for outages, receivers not being in the water, or any other reason why a receiver did not detect any or as many fish as we would expect.

First, let's load the required packages and data.

```{r}
library(rgeos)
library(maptools)
library(klesdatr)
library(magrittr)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)

receiver <- klesdatr::receiver %>% 
  mutate(Section_n = as.integer(Section),
         Receiver_n = as.integer(Receiver))
glimpse(receiver)

#fish_with_location <- klesdatr::calculate_location() %>%
#  mutate(Section_n = as.integer(Section))
#print(glimpse(fish_with_location))

detection <- klesdatr::detection %>% 
  mutate(Receiver_n = as.integer(Receiver))
glimpse(detection)

section <- klesdatr::section %>%
  mutate(Section_n = as.integer(Section))
glimpse(section)
```


We have `r detection$Receiver_n %>% unique() %>% length()` distinct receivers, but let's start by looking at aggregated statistics.


## Aggregate statistics

First we plot the total number of sightings by section. We copy code from `klesdatr::plot_section()` and aggregate the detection dates by section.

```{r}
section_with_detections <- section %>% 
  dplyr::right_join(detection %>%
                      join(receiver) %>% 
                      group_by(Section_n) %>% 
                      summarize(total_sightings = n()) %>% 
                      mutate(fraction_total_sightings = total_sightings/sum(total_sightings))
                    )

if (requireNamespace("rgeos", quietly = TRUE) & requireNamespace("maptools", quietly = TRUE)) {
    section_sp  <- klesdatr::section_sp %>% ggplot2::fortify(region = "Section")
    poly <- list(ggplot2::geom_polygon(data = dplyr::filter_(section_sp, ~!hole),
                                       ggplot2::aes_string(x = "long / 1000", y = "lat / 1000", group = "id"),
                                       alpha = 1/5,
                                       color = "grey50"),
                 ggplot2::geom_polygon(data = dplyr::filter_(section_sp, ~!hole),
                                       ggplot2::aes_string(x = "long / 1000", y = "lat / 1000", group = "id"),
                                       alpha = 1/5,
                                       color = "grey50"))
    } else {
    message("The packages 'maptools' and rgeos' are required for the full version of this spatial plot.")
    poly <- NULL
    }

ggplot2::ggplot(data = section_with_detections,
                ggplot2::aes_string(x = "SectionX / 1000", y = "SectionY / 1000")) +
    ggplot2::geom_point(aes(size = fraction_total_sightings)) +
    poly + 
    ggplot2::geom_text(aes(x = SectionX / 1000 + 2.5, label = Section_n), color = "blue", nudge_x = 2, size = 6) +
    ggplot2::coord_equal() +
    ggplot2::scale_x_continuous(name = "Easting (km)") +
    ggplot2::scale_y_continuous(name = "Northing (km)")
```


Looking at the `DetectionDate` we see that the number of sightings is nowhere near constant...

```{r}
detection %>% 
  group_by(DetectionDate) %>% 
  summarize(total_sightings = n()) %>% 
  ggplot(aes(x = DetectionDate, y = total_sightings, label = total_sightings)) +
    geom_density(stat = "identity")
```

...but there don't seem to be obvious seasonal patterns.

```{r}
detection %>% 
  mutate(month = lubridate::month(DetectionDate)) %>% 
  group_by(month) %>% 
  summarize(total_detections = n()) %>% 
  ggplot(aes(x = month, y = total_detections, label = month.abb[month])) +
    geom_bar(stat = "identity") +
    geom_text(nudge_y = 80)
```  

[Todo: Add in how many receivers were active]


## Individual receivers

Looking at the number of sightings per receiver we see huge differences. Indeed, some receivers rarely picked up any signals.

```{r}
detection %>% 
  group_by(Receiver_n) %>% 
  summarize(total_sightings = n()) %>% 
  ggplot(aes(x = Receiver_n, y = total_sightings, label = total_sightings)) +
    geom_bar(stat = "identity") +
    geom_text(color = "red", vjust = 0)
```

Here is a zoom at all the sightings for each receiver

```{r}
detection %>% 
#  filter(Receiver_n == 1) %>% 
  ggplot(aes(x = DetectionDate)) + 
    geom_point(aes(y = 0), shape = 124) +
    geom_density(kernel = "epanechnikov") +
    facet_wrap(~ Receiver, scales = "free_y")
```


[Todo: Highlight gaps in detection. Did other receivers in the same section get any signal during those gaps?]
