---
title: "Analysis of Coverage"
author: "Bernhard Konrad"
date: "December 22, 2015"
output:
  html_document:
    keep_md: true
---

In this document we run a data quality analysis to check the observed receiver coverage. This is to control for outages, receivers not being in the water, or any other reason why a receiver did not detect any or as many fish as we would expect.

First, let's load the required packages and data.

```{r}
library(rgeos)
library(maptools)
library(klesdatr)
library(magrittr)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)

receiver <- klesdatr::receiver %>% 
  mutate(Section_n = as.integer(Section),
         Receiver_n = as.integer(Receiver))
glimpse(receiver)

#fish_with_location <- klesdatr::calculate_location() %>%
#  mutate(Section_n = as.integer(Section))
#print(glimpse(fish_with_location))

detection <- klesdatr::detection %>% 
  mutate(Receiver_n = as.integer(Receiver))
glimpse(detection)

section <- klesdatr::section %>%
  mutate(Section_n = as.integer(Section))
glimpse(section)
```


We have `r detection$Receiver_n %>% unique() %>% length()` distinct receivers, but let's start by looking at aggregated statistics.


## Aggregate statistics

First we plot the total number of sightings by section. We copy code from `klesdatr::plot_section()` and aggregate the detection dates by section.

```{r}
section_with_detections <- section %>% 
  dplyr::right_join(detection %>%
                      join(receiver) %>% 
                      group_by(Section_n) %>% 
                      summarize(total_sightings = n()) %>% 
                      mutate(fraction_total_sightings = total_sightings/sum(total_sightings))
                    )

if (requireNamespace("rgeos", quietly = TRUE) & requireNamespace("maptools", quietly = TRUE)) {
    section_sp  <- klesdatr::section_sp %>% ggplot2::fortify(region = "Section")
    poly <- list(ggplot2::geom_polygon(data = dplyr::filter_(section_sp, ~!hole),
                                       ggplot2::aes_string(x = "long / 1000", y = "lat / 1000", group = "id"),
                                       alpha = 1/5,
                                       color = "grey50"),
                 ggplot2::geom_polygon(data = dplyr::filter_(section_sp, ~!hole),
                                       ggplot2::aes_string(x = "long / 1000", y = "lat / 1000", group = "id"),
                                       alpha = 1/5,
                                       color = "grey50"))
    } else {
    message("The packages 'maptools' and rgeos' are required for the full version of this spatial plot.")
    poly <- NULL
    }

ggplot2::ggplot(data = section_with_detections,
                ggplot2::aes_string(x = "SectionX / 1000", y = "SectionY / 1000")) +
    ggplot2::geom_point(aes(size = fraction_total_sightings)) +
    poly + 
    ggplot2::geom_text(aes(x = SectionX / 1000 + 2.5, label = Section_n), color = "blue", nudge_x = 2, size = 6) +
    ggplot2::coord_equal() +
    ggplot2::scale_x_continuous(name = "Easting (km)") +
    ggplot2::scale_y_continuous(name = "Northing (km)")
```


Looking at the `DetectionDate` we see that the number of sightings is nowhere near constant...

```{r}
deployment %>% 
  group_by(DeploymentDate) %>% 
  summarize(active_receivers = n()) %>% 
  left_join(detection %>% 
              group_by(DetectionDate) %>% 
              summarize(total_sightings = n()),
            by = c("DeploymentDate" = "DetectionDate")
            ) %>% 
  ggplot() +
    geom_path(aes(x = DeploymentDate, y = total_sightings)) +
    geom_path(aes(x = DeploymentDate, y = active_receivers), color = "red") + 
    geom_text(aes(x = as.Date(ymd("2014-01-01")), y = 36), label = "active receivers", color = "red", size = 7)
  
```

...but there don't seem to be obvious seasonal patterns.

```{r}
deployment %>% 
  group_by(DeploymentDate) %>% 
  summarize(active_receivers = n()) %>% 
  left_join(detection, by = c("DeploymentDate" = "DetectionDate")) %>% 
  mutate(month = lubridate::month(DeploymentDate)) %>% 
  group_by(month) %>% 
  summarize(total_detections = n(),
            monthly_active_receivers = mean(active_receivers)
            ) %>% 
  mutate(normalized_total_detections = total_detections/sum(total_detections),
            normalized_monthly_active_receivers = monthly_active_receivers/sum(monthly_active_receivers)) %>% 
  ggplot(aes(x = month, y = normalized_total_detections, label = month.abb[month])) +
    geom_bar(stat = "identity") +
    geom_line(aes(y = normalized_monthly_active_receivers), color = "red", size = 2) +
    geom_text(aes(x = 9, y = 0.095), label = "normalized mean \n # active receivers", color = "red", size = 5) +
  geom_text(nudge_y = 0.002)
```  


## Individual receivers

Looking at the number of sightings per receiver we see huge differences. Indeed, some receivers rarely picked up any signals.

```{r}
detection %>% 
  group_by(Receiver_n) %>% 
  summarize(total_sightings = n()) %>% 
  ggplot(aes(x = Receiver_n, y = total_sightings, label = total_sightings)) +
    geom_bar(stat = "identity") +
    geom_text(color = "red", vjust = 0)
```

Here is a zoom at all the sightings for each receiver

```{r}
detection %>% 
#  filter(Receiver_n == 1) %>% 
  ggplot(aes(x = DetectionDate)) + 
    geom_point(aes(y = 0), shape = 124) +
    geom_density(kernel = "epanechnikov") +
    facet_wrap(~ Receiver, scales = "free_y")
```


[Todo: Highlight gaps in detection. Did other receivers in the same section get any signal during those gaps?]


```{r}
# Folds in the non-active days into deployment. May not be as useful as I thought, but now it's done and can be used if convenient.

df_date_sequence <- seq(min(deployment$DeploymentDate), max(deployment$DeploymentDate), by = '1 day')

non_active_days <- function(n_receiver){
  setdiff(ymd(as.character(df_date_sequence)), ymd(as.character(deployment$DeploymentDate[as.integer(deployment$Receiver) == n_receiver]))) %>% 
    as.POSIXct(origin="1970-01-01") %>% 
    as.Date
}

df_with_nonActive_days <- function(n_receiver){
  if (length(non_active_days(n_receiver)) > 0) {
    data.frame(
      dayte = non_active_days(n_receiver),
      Receiver_n = n_receiver,
      was_active = FALSE
    )
  } else NULL
}

complete_df <- function(){
  df <- deployment %>% 
    transmute(
      dayte = DeploymentDate,
      was_active = TRUE,
      Receiver_n = as.integer(Receiver))
  for (n in unique(df$Receiver_n)) {
    new <- df_with_nonActive_days(n)
    if (!is.null(new)) df <- rbind(df, new)
  }
  df
}

full_deployment <- complete_df()

stopifnot(length(df_date_sequence) * length(unique(deployment$Receiver)) == nrow(full_deployment))
```
